State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-04-15 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
small updates

The state estimation step utilises the Unscented Rauch Tung Striebel Smoother (URTSS)~\cite{Sarkka2010} to determine the posterior state sequence. The (URTSS) uses the Unscented Kalman Filter (UKF) in the forward iteration~\cite{Julier1997} to calculate a forward posterior state estimate, $\hat{\mathbf x}_t^{f}$, followed by a backward path to compute the smoothed state estimations $\hat{\mathbf x}_t^{b}$. The method is appropriate for state estimation of stochastic nonlinear dynamical systems. Alternatives to the UKF are the Extended Kalman Filter (EKF) \cite{Haykin2001}  and Sequential Monte Carlo (SMC) filtering \cite{doucet2001}. The extended Kalman filter approximates the state transition equation by linearizing about the fixed points (by calculating the Jacobian) at each time step. The linearization maintains the Gaussianality in the model. The unscented Kalman filter approximates the \emph{a posteriori} state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, whilst maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, as the EKF maintains a first order approximation where the UKF provides an approximation accurate at least to the second order. In addition, the computation of the Jacobian used in the
EKF can be problematic. SMC filtering can theoretically provide an exact \emph{a posteriori} state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to computation demands.

>>>>

The state estimation step utilises the Unscented Rauch Tung Striebel Smoother (URTSS)~\cite{Sarkka2010} to determine the posterior state sequence. Alternatives to the UKF are the Extended Kalman Filter (EKF)~\cite{Haykin2001} and Sequential Monte Carlo (SMC) filtering \cite{doucet2001}. The extended Kalman filter approximates the state transition equation by linearizing about the fixed points (by calculating the Jacobian) at each time instance. The linearisation maintains the Gaussianality of the model. The unscented Kalman filter approximates the \emph{a posteriori} state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, whilst maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, as the EKF maintains a first order approximation where the UKF provides an approximation accurate at least to the second order. In addition, the computation of the Jacobian used in the
EKF can be problematic. SMC filtering can theoretically provide an exact \emph{a posteriori} state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to computation demands.The (URTSS) uses the Unscented Kalman Filter (UKF) in the forward iteration~\cite{Julier1997} to calculate a forward posterior state estimate, $\hat{\mathbf x}_t^{f}$, followed by a backward path to compute the smoothed state estimations $\hat{\mathbf x}_t^{b}$. The method is appropriate for state estimation of stochastic nonlinear dynamical systems.
